---
title: "COVID19 real-time analysis: NHS Pathways data importing & cleaning"
author: "Quentin J. Leclerc, Thibaut Jombart"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 4
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 6,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE)
```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
`.R` files contained in `scripts` at the root of the factory are automatically
loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.



## Load packages

```{r libraries}

library(reportfactory)
library(here)
library(tidyverse)
library(linelist)
library(xml2)
library(rvest)

```



## Load scripts

These scripts will load:

* all local scripts, stored as `.R` filesinside `/scripts/`
* all global scripts, i.e. stored outside the factory in `../scripts/`

```{r load_scripts}

rfh_load_scripts()

```





## Load data

The latest NHS Pathways data is automatically detected using
`fetch_url_pathways_calls()` and `fetch_url_pathways_online()`. We also recover the date from the url:

```{r load_data}

#fetch the latest data
current_pathways_calls <- fetch_url_pathways_calls() 
current_pathways_online <- fetch_url_pathways_online()

#get the date of the latest data
pathways_calls_date <- linelist::guess_dates(current_pathways_calls)
pathways_online_date <- linelist::guess_dates(current_pathways_online)

```

We load the data:

```{r pathways_import}

#import the data
pathways_calls <- rio::import(current_pathways_calls)
pathways_online <- rio::import(current_pathways_online)

```

The **completion date** of the NHS Pathways 111 and 999 calls data is
**`r format(pathways_calls_date, format = "%A %d %b %Y")`**, and the 
**completion date** of the NHS Pathways 111 online data is
**`r format(pathways_online_date, format = "%A %d %b %Y")`**.






<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Process data {.tabset .tabset-fade .tabset-pills}

## Clean data

We first clean the data:

```{r clean_data_calls}

pathways_calls <- linelist::clean_data(pathways_calls,
                                       wordlists = cleaning_rules,
                                       guess_dates = FALSE)
pathways_calls$call_date <- linelist::guess_dates(pathways_calls$call_date, error_tolerance = 1)
head(pathways_calls)

```

```{r clean_data_online}

pathways_online <- linelist::clean_data(pathways_online,
                                        wordlists = cleaning_rules,
                                        guess_dates = FALSE)
pathways_online$journeydate <- linelist::guess_dates(pathways_online$journeydate, error_tolerance = 1)
head(pathways_calls)

```

## Rename data

We then rename the variables in the data:

```{r rename_variables}

pathways_calls <- pathways_calls %>%
  rename(site_type = contains("sitetype"),
         date = contains("call_date"),
         age = contains("ageband"),
         ccg_code = contains("ccgcode"),
         ccg_name = contains("ccgname"),
         count = contains("triagecount"))

pathways_online <- pathways_online %>%
  rename(date = contains("journeydate"),
         age = contains("ageband"),
         ccg_code = contains("ccgcode"),
         ccg_name = contains("ccgname"),
         count = contains("total"))

#small edit for consistency in age groups
pathways_online$age <- gsub("70_years", "70_120_years", pathways_online$age)

```

## Merge both datasets

We merge the 111 and 999 calls with the 111 online data:

```{r merge_datasets}

site_type <- rep("111_online", nrow(pathways_online))

pathways_online <- cbind(site_type, pathways_online)

pathways_all <- rbind(pathways_calls, pathways_online)

```


## Reformat age classes

Finally, we correct the formatting of the age classes:

```{r reformat_ages}

pathways_all$age <- gsub("_years", "", pathways_all$age)
pathways_all$age <- gsub("_", "-", pathways_all$age)
head(pathways_all)

```

## Add variables

We also add NHS regions to the data:

```{r add_nhs}

ccg_info <- rio::import(here::here("data", "csv", "ccg_info_april_2020.csv"))

ccg_info <- ccg_info %>%
  as_tibble() %>%
  clean_data()

head(ccg_info)

pathways_all <- left_join(pathways_all,
                          rename(ccg_info, ccg_name = name))
head(pathways_all)

```


## Add timestamp

Add dataset version date as an attribute for later reference.

```{r add_timestamp}

attr(pathways_all, "timestamp") <- pathways_calls_date

```





<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Export data {.tabset .tabset-fade .tabset-pills}

## R objects

We export the clear database, placed in `produced_rds/` as well as in
`data/rds`:

```{r export_rds}

## check if a directory exists and if not then creates it
if (!dir.exists("produced_rds")) {
  dir.create("produced_rds")
}

## create the text for the file name with the database date
rds_file_name <- sprintf("pathways_%s.rds",
                         format(pathways_calls_date, "%Y-%m-%d"))
rds_file_name

## save the rds file in the produced_rds folder
saveRDS(pathways_all,
        file.path("produced_rds", rds_file_name))
```

We copy these files to the `data/rds` folder:

```{r copy_rds}
## copy some files into `data/rds`

## provide the destination of where to copy the data
destination <- here("data",
                    "rds",
                    rds_file_name)
## copy the rds data
file.copy(from = file.path("produced_rds", rds_file_name),
          to = destination,
          overwrite = TRUE)

## copy to the generic 'latest' data file
destination <- here("data",
                    "rds",
                    "pathways_latest.rds")
file.copy(from = file.path("produced_rds", rds_file_name),
          to = destination,
          overwrite = TRUE)

```




## Update the `current_pathways.R` script

```{r update_script}

## path to the output file 
script_destination <- here::here("scripts",
                                 "current_pathways.R")

## comments to say when this was updated
txt <- paste("## This file is generated automatically by `clean_latest_pathways`",
             "## Do not edit it by hand!\n",
             sep = "\n")
cat(txt, file = script_destination, append = FALSE)

txt <- sprintf("\n## This file was last updated on the: %s\n",
               Sys.time())
cat(txt, file = script_destination, append = TRUE)

## actual content of the script
txt <- sprintf('\ncurrent_pathways <- here::here("data",
                                 "rds",
                                 "%s")',
               rds_file_name)
cat(txt, file = script_destination, append = TRUE)

```







<!-- =======================================================  -->
<!-- =======================================================  -->
<!-- ======================================================= -->

# System information {.tabset .tabset-fade .tabset-pills}

## Outline

The following information documents the system on which the document was
compiled.


## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```
