---
title: "Analysis of 111/999 NHS calls and 111 online reports"
author: "Quentin Leclerc, Emily Nightingale, Samuel Abbott & Thibaut Jombart, for CMMID"
date: "`r format(Sys.time(), '%A %d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: zenburn
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_collapse: no
    toc_depth: 4
    toc_float: yes
    css: !expr here::here('css', 'style.css')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      collapse = TRUE,
                      fig.width = 8,
                      fig.height = 4,
                      dpi = 150,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "figures/nhs_paths/")
```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->

# Data preparation {.tabset .tabset-fade .tabset-pills}

## Outline


* **Load scripts**: loads libraries and useful scripts used in the analyses; all
`.R` files contained in `scripts` at the root of the factory are automatically
loaded

* **Load data**: imports datasets, and may contain some *ad hoc* changes to the
data such as specific data cleaning (not used in other reports), new variables
used in the analyses, etc.



## Load packages

```{r libraries}

library(reportfactory)
library(here)
library(rio) 
library(tidyverse)
library(incidence)
library(distcrete)
library(epitrix)
library(earlyR)
library(projections)
library(linelist)
library(remotes)
library(janitor)
library(kableExtra)
library(DT)
library(cyphr)
library(chngpt)
library(lubridate)
library(ggpubr)

```



## Load scripts

These scripts will load:

* all scripts stored as `.R` files inside `/scripts/`
* all scripts stored as `.R` files inside `/src/`

These scripts also contain routines to access the latest clean encrypted data
(see next section). 

```{r load_scripts}

reportfactory::rfh_load_scripts()

```





## Load clean data

We import the latest NHS pathways data:


```{r load_data}

x <- import_pathways() %>%
  as_tibble()
x

```

We also import demographics data for NHS regions in England, used later in our analysis:

```{r nhs_region_pop}

path <- here::here("data", "csv", "nhs_region_population_2018.csv")
nhs_region_pop <- rio::import(path) %>%
  mutate(nhs_region = str_to_title(gsub("_"," ",nhs_region)))

nhs_region_pop$nhs_region <- gsub(" Of ", " of ", nhs_region_pop$nhs_region)
nhs_region_pop$nhs_region <- gsub(" And ", " and ", nhs_region_pop$nhs_region)
nhs_region_pop

```

Finally, we import publically available deaths per NHS region:

```{r import_deaths}

dth <- import_deaths() %>%
  mutate(nhs_region = str_to_title(gsub("_"," ",nhs_region)))

dth$nhs_region <- gsub(" Of ", " of ", dth$nhs_region)
dth$nhs_region <- gsub(" And ", " and ", dth$nhs_region)
dth

```

## Completion date

We extract the completion date from the NHS Pathways file timestamp:

```{r database_date}

database_date <- attr(x, "timestamp")
database_date

```

The **completion date** of the NHS Pathways data is
**`r format(database_date, format = "%A %d %b %Y")`**.




## Add variables

We add the following variable:

* `day`: an integer representing the number of days from the earliest data
reported, used for modelling purposes; the first day is 0

```{r add_variables}

x <- x %>% 
  mutate(nhs_region = str_to_title(gsub("_"," ",nhs_region)),
         nhs_region = gsub(" Of ", " of ", nhs_region),
         nhs_region = gsub(" And ", " and ", nhs_region),
         day = as.integer(date - min(date, na.rm = TRUE)))

```



## Auxiliary functions

These are functions which will be used further in the analyses.

Function to estimate the generalised R-squared as the proportion of deviance
explained by a given model:

```{r Rsq}

## Function to calculate R2 for Poisson model
## not adjusted for model complexity but all models have the same DF here

Rsq <- function(x) {
  1 - (x$deviance / x$null.deviance)
}

```

Function to extract growth rates per region as well as halving times, and the
associated 95% confidence intervals:

```{r get_r}

## function to extract the coefficients, find the level of the intercept,
## reconstruct the values of r, get confidence intervals

get_r <- function(model) {
  ##  extract coefficients and conf int
  out <- data.frame(r = coef(model))  %>%
    rownames_to_column("var") %>% 
    cbind(confint(model)) %>%
    filter(!grepl("day_of_week", var)) %>% 
    filter(grepl("day", var)) %>%
    rename(lower_95 = "2.5 %",
           upper_95 = "97.5 %") %>%
    mutate(var = sub("day:", "", var))
  
  ## reconstruct values: intercept + region-coefficient
  for (i in 2:nrow(out)) {
    out[i, -1] <- out[1, -1] + out[i, -1]
  }
  
  ## find the name of the intercept, restore regions names
  out <- out %>%
    mutate(nhs_region = model$xlevels$nhs_region) %>%
    select(nhs_region, everything(), -var)
  
  ## find halving times
  halving <- log(0.5) / out[,-1] %>%
    rename(halving_t = r,
           halving_t_lower_95 = lower_95,
           halving_t_upper_95 = upper_95)
  
  ## set halving times with exclusion intervals to NA
  no_halving <- out$lower_95 < 0 & out$upper_95 > 0
  halving[no_halving, ] <- NA_real_
  
  ## return all data
  cbind(out, halving)
  
}

```

Function to classify the day of the week into *weekend*, *Monday*, and *the rest*:

```{r day_of_week}

## Fn to add day of week
day_of_week <- function(df) {
  df %>% 
    dplyr::mutate(day_of_week = lubridate::wday(date, label = TRUE)) %>% 
    dplyr::mutate(day_of_week = dplyr::case_when(
      day_of_week %in% c("Sat", "Sun") ~ "weekend",
      day_of_week %in% c("Mon") ~ "monday",
      !(day_of_week %in% c("Sat", "Sun", "Mon")) ~ "rest_of_week"
    ) %>% 
      factor(levels = c("rest_of_week", "monday", "weekend")))
}

```

Custom color palettes, color scales, and vectors of colors:

```{r palettes}

pal <- c("#006212",
         "#ae3cab",
         "#00db90",
         "#960c00",
         "#55aaff",
         "#ff7e78",
         "#00388d")

blue_pal <- c("#8cb5f5","#00a1bd","#004480")

```




<!-- ======================================================== -->
<!-- ======================================================== -->
<!-- ======================================================== -->

# Spatio-temporal trends in NHS Pathways reports {.tabset .tabset-fade .tabset-pills}

## Outline

We look for spatio-temporal patterns in COVID-19 related 111/999 calls and 111 
online reports. Analyses are broken down by NHS region. We also look for estimates
of recent growth rate and associated doubling / halving time.

## Reports by region over time

```{r all_volume, fig.height = 6}

tab_date_region_all <- x %>%
  filter(!is.na(nhs_region)) %>%
  group_by(date, nhs_region) %>%
  summarise(n = sum(count))

tab_date_region_all %>%
  ggplot(aes(x = date, y = n, color = nhs_region)) +
  geom_line() +
  geom_point() +
  geom_smooth(method = "loess", span = .5, color = "black") +
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "dashed") +
  theme_bw() +
  large_txt +
  scale_weeks +
  theme(legend.position = "bottom") +
  rotate_x + 
  guides(color = guide_legend(title = "", ncol = 3)) +
  labs(x = NULL,
       y = "Number of daily reports") +
  scale_colour_manual(values = pal)

tab_date_region_all %>% 
  adorn_totals() %>%
  spread(nhs_region, n, fill = 0, drop = FALSE) %>%
  adorn_totals(c("row", "col")) %>% 
  dt_tab()

```


## Reports by region by age over time

We plot the number of 111/999 calls and 111 online reports by age, and the 
proportion of 111/999 calls and 111 online reports by age.
In the second graph, the vertical lines indicate the proportion of individuals 
residing in the corresponding NHS region who belong to the corresponding age group. 

```{r all_volume_age, fig.height = 6}

tab_date_region_age_all <- x %>%
  filter(!is.na(nhs_region),
         age != "missing") %>%
  group_by(date, nhs_region, age) %>%
  summarise(n = sum(count))

tab_date_region_age_all %>%
  ggplot(aes(x = date, y = n, fill = age)) +
  geom_col(position = "stack") +
  scale_fill_manual(values = blue_pal) +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(fill = guide_legend(title = "Age", ncol = 3)) +
  labs(x = NULL,
       y = "Total daily reports by age") +
  facet_wrap(~ nhs_region, ncol = 4)


tab_date_region_age_all <- tab_date_region_age_all %>%
  group_by(date, nhs_region) %>%
  summarise(tot = sum(n)) %>%
  left_join(tab_date_region_age_all, by = c("date", "nhs_region")) %>%
  mutate(prop_n = n/tot)

tab_date_region_age_all %>%
  ggplot(aes(x = date, y = prop_n, color = age)) +
  scale_color_manual(values = blue_pal) +
  geom_line() +
  geom_point() +
  geom_hline(data = nhs_region_pop, aes(yintercept = value, color = variable)) +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  guides(color = guide_legend(title = "Age", ncol = 3)) +
  labs(x = NULL,
       y = "Proportion of daily reports by age") +
  facet_wrap(~ nhs_region, ncol = 4)
```


## Moving window model

We fit quasi-Poisson GLMs for 14-day windows to get growth rates over time.
```{r sliding_model}
## set moving time window (1/2/3 weeks)
w <- 14

# create empty df
r_all_sliding <- NULL

## make data for model
x_model_all_moving <- x %>%
  filter(!is.na(nhs_region)) %>% 
  group_by(date, nhs_region) %>%
  summarise(n = sum(count)) 

unique_dates <- unique(x_model_all_moving$date)

for (i in 1:(length(unique_dates) - w)) {
  
  date_i <- unique_dates[i]
  
  date_i_max <- date_i + w
  
  model_data <- x_model_all_moving %>%
    filter(date >= date_i & date < date_i_max) %>%
    mutate(day = as.integer(date - date_i)) %>% 
    day_of_week()
  
  
  mod <- glm(n ~ day * nhs_region + day_of_week,
             data = model_data,
             family = 'quasipoisson')
  
  # get growth rate
  r <- get_r(mod)
  r$w_min <- date_i
  r$w_max <- date_i_max
  
  # combine all estimates
  r_all_sliding <- bind_rows(r_all_sliding, r)
  
}

#serial interval distribution
SI_param = epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale,
                                        w = 0.5)

#convert growth rates r to R0
r_all_sliding <- r_all_sliding %>%
  mutate(R = epitrix::r2R0(r, SI_distribution),
         R_lower_95 = epitrix::r2R0(lower_95, SI_distribution),
         R_upper_95 = epitrix::r2R0(upper_95, SI_distribution))

```


We examine the evolution of the growth rate by region over time.

```{r all_sliding_growth}
# plot
plot_growth <-
  r_all_sliding %>%
  ggplot(aes(x = w_max, y = r)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(colour = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated daily growth rate (r)") +
  scale_colour_manual(values = pal)

```

From the growth rate, we derive R and examine its value through time.

```{r all_sliding_R}
# plot
plot_R <-
  r_all_sliding %>%
  ggplot(aes(x = w_max, y = R)) +
  geom_ribbon(aes(ymin = R_lower_95, ymax = R_upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated effective reproduction\nnumber (Re)") +
  scale_colour_manual(values = pal)

```

```{r fig2_3}

R <- r_all_sliding %>%
  mutate(lower_95 = R_lower_95, 
         upper_95 = R_upper_95,
         value = R,
         measure = "R",
         reference = 1)

r_R <- r_all_sliding %>%
  mutate(measure = "r",
         value = r,
         reference = 0) %>%
  bind_rows(R)

r_R %>%
  ggplot(aes(x = w_max, y = value)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(aes(yintercept = reference), linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0,0, "cm"),
        strip.background = element_blank(),
        # strip.text.x = element_blank(),
        strip.placement = "outside"
  ) +
  guides(color = guide_legend(title = "",
                              override.aes = list(fill = NA)),
         fill = FALSE) +
  labs(x = "", y = "") +
  scale_colour_manual(values = pal) +
  facet_grid(rows = vars(measure),
             scales = "free_y",
             switch = "y",
             labeller = as_labeller(c(r = "Daily growth rate (r)",
                                      R = "Effective reproduction\nnumber (Re)")))



```



## Moving window model - stratified by age

We repeat the above analysis, where we fit quasi-Poisson GLMs for 14-day windows 
to get growth rates over time, but apply this to each age group separately (0-18,
19-69, 70-120 years old).

We first run the analysis for 0-18 years old.

```{r sliding_model_0_18}
## set moving time window (2 weeks)
w <- 14

# create empty df
r_all_sliding_0_18 <- NULL

## make data for model
x_model_all_moving_0_18 <- x %>%
  filter(!is.na(nhs_region),
         age == "0-18") %>% 
  group_by(date, nhs_region) %>%
  summarise(n = sum(count)) 

unique_dates <- unique(x_model_all_moving_0_18$date)

for (i in 1:(length(unique_dates) - w)) {
  
  date_i <- unique_dates[i]
  
  date_i_max <- date_i + w
  
  model_data <- x_model_all_moving_0_18 %>%
    filter(date >= date_i & date < date_i_max) %>%
    mutate(day = as.integer(date - date_i)) %>% 
    day_of_week()
  
  
  mod <- glm(n ~ day * nhs_region + day_of_week,
             data = model_data,
             family = 'quasipoisson')
  
  # get growth rate
  r <- get_r(mod)
  r$w_min <- date_i
  r$w_max <- date_i_max
  
  # combine all estimates
  r_all_sliding_0_18 <- bind_rows(r_all_sliding_0_18, r)
  
}

#serial interval distribution
SI_param = epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale, w = 0.5)

#convert growth rates r to R0
r_all_sliding_0_18 <- r_all_sliding_0_18 %>%
  mutate(R = epitrix::r2R0(r, SI_distribution),
         R_lower_95 = epitrix::r2R0(lower_95, SI_distribution),
         R_upper_95 = epitrix::r2R0(upper_95, SI_distribution))

```

```{r all_sliding_growth_0_18}
# plot
plot_growth <-
  r_all_sliding_0_18 %>%
  ggplot(aes(x = w_max, y = r)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(colour = guide_legend(title = "",override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated daily growth rate (r)"
  ) +
  scale_colour_manual(values = pal)

```

```{r all_sliding_R_0_18}
# plot
plot_R <-
  r_all_sliding_0_18 %>%
  ggplot(aes(x = w_max, y = R)) +
  geom_ribbon(aes(ymin = R_lower_95, ymax = R_upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated effective reproduction\nnumber (Re)"
  ) +
  scale_colour_manual(values = pal)

```


```{r fig2_3_0_18}

R <- r_all_sliding_0_18 %>%
  mutate(lower_95 = R_lower_95, 
         upper_95 = R_upper_95,
         value = R,
         measure = "R",
         reference = 1)

r_R <- r_all_sliding_0_18 %>%
  mutate(measure = "r",
         value = r,
         reference = 0) %>%
  bind_rows(R)

fig2_3_0_18 <- r_R %>%
  ggplot(aes(x = w_max, y = value)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(aes(yintercept = reference), linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0,0, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside"
  ) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "", y = "") +
  scale_colour_manual(values = pal) +
  facet_grid(rows = vars(measure),
             scales = "free_y",
             switch = "y",
             labeller = as_labeller(c(r = "Daily growth rate (r)",
                                      R = "Effective reproduction\nnumber (Re)")))

```

Then, we run the analysis for 19-69 years old.

```{r sliding_model_19_69}
## set moving time window (2 weeks)
w <- 14

# create empty df
r_all_sliding_19_69 <- NULL

## make data for model
x_model_all_moving_19_69 <- x %>%
  filter(!is.na(nhs_region),
         age == "19-69") %>% 
  group_by(date, nhs_region) %>%
  summarise(n = sum(count)) 

unique_dates <- unique(x_model_all_moving_19_69$date)

for (i in 1:(length(unique_dates) - w)) {
  
  date_i <- unique_dates[i]
  
  date_i_max <- date_i + w
  
  model_data <- x_model_all_moving_19_69 %>%
    filter(date >= date_i & date < date_i_max) %>%
    mutate(day = as.integer(date - date_i)) %>% 
    day_of_week()
  
  
  mod <- glm(n ~ day * nhs_region + day_of_week,
             data = model_data,
             family = 'quasipoisson')
  
  # get growth rate
  r <- get_r(mod)
  r$w_min <- date_i
  r$w_max <- date_i_max
  
  # combine all estimates
  r_all_sliding_19_69 <- bind_rows(r_all_sliding_19_69, r)
  
}

#serial interval distribution
SI_param = epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale, w = 0.5)

#convert growth rates r to R0
r_all_sliding_19_69 <- r_all_sliding_19_69 %>%
  mutate(R = epitrix::r2R0(r, SI_distribution),
         R_lower_95 = epitrix::r2R0(lower_95, SI_distribution),
         R_upper_95 = epitrix::r2R0(upper_95, SI_distribution))

```

```{r all_sliding_growth_19_69}
# plot
plot_growth <-
  r_all_sliding_19_69 %>%
  ggplot(aes(x = w_max, y = r)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(colour = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated daily growth rate (r)") +
  scale_colour_manual(values = pal)

```

```{r all_sliding_R_19_69}
# plot
plot_R <-
  r_all_sliding_19_69 %>%
  ggplot(aes(x = w_max, y = R)) +
  geom_ribbon(aes(ymin = R_lower_95, ymax = R_upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated effective reproduction\nnumber (Re)"
  ) +
  scale_colour_manual(values = pal)

```

```{r fig2_3_19_69}

R <- r_all_sliding_19_69 %>%
  mutate(lower_95 = R_lower_95, 
         upper_95 = R_upper_95,
         value = R,
         measure = "R",
         reference = 1)

r_R <- r_all_sliding_19_69 %>%
  mutate(measure = "r",
         value = r,
         reference = 0) %>%
  bind_rows(R)

fig2_3_19_69 <- r_R %>%
  ggplot(aes(x = w_max, y = value)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(aes(yintercept = reference), linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0,0, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside"
  ) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "", y = "") +
  scale_colour_manual(values = pal) +
  facet_grid(rows = vars(measure),
             scales = "free_y",
             switch = "y",
             labeller = as_labeller(c(r = "Daily growth rate (r)",
                                      R = "Effective reproduction\nnumber (Re)")))

```

Finally, we run the analysis for 70-120 years old.

```{r sliding_model_70_120}
## set moving time window (2 weeks)
w <- 14

# create empty df
r_all_sliding_70_120 <- NULL

## make data for model
x_model_all_moving_70_120 <- x %>%
  filter(!is.na(nhs_region),
         age == "70-120") %>% 
  group_by(date, nhs_region) %>%
  summarise(n = sum(count)) 

unique_dates <- unique(x_model_all_moving_70_120$date)

for (i in 1:(length(unique_dates) - w)) {
  
  date_i <- unique_dates[i]
  
  date_i_max <- date_i + w
  
  model_data <- x_model_all_moving_70_120 %>%
    filter(date >= date_i & date < date_i_max) %>%
    mutate(day = as.integer(date - date_i)) %>% 
    day_of_week()
  
  
  mod <- glm(n ~ day * nhs_region + day_of_week,
             data = model_data,
             family = 'quasipoisson')
  
  # get growth rate
  r <- get_r(mod)
  r$w_min <- date_i
  r$w_max <- date_i_max
  
  # combine all estimates
  r_all_sliding_70_120 <- bind_rows(r_all_sliding_70_120, r)
  
}

#serial interval distribution
SI_param = epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale, w = 0.5)

#convert growth rates r to R0
r_all_sliding_70_120 <- r_all_sliding_70_120 %>%
  mutate(R = epitrix::r2R0(r, SI_distribution),
         R_lower_95 = epitrix::r2R0(lower_95, SI_distribution),
         R_upper_95 = epitrix::r2R0(upper_95, SI_distribution))

```

```{r all_sliding_growth_70_120}
# plot
plot_growth <-
  r_all_sliding_70_120 %>%
  ggplot(aes(x = w_max, y = r)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(colour = guide_legend(title = "",override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated daily growth rate (r)"
  ) +
  scale_colour_manual(values = pal)

```

```{r all_sliding_R_70_120}
# plot
plot_R <-
  r_all_sliding_70_120 %>%
  ggplot(aes(x = w_max, y = R)) +
  geom_ribbon(aes(ymin = R_lower_95, ymax = R_upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated effective reproduction\nnumber (Re)") +
  scale_colour_manual(values = pal)

```

```{r fig2_3_70_120}

R <- r_all_sliding_70_120 %>%
  mutate(lower_95 = R_lower_95, 
         upper_95 = R_upper_95,
         value = R,
         measure = "R",
         reference = 1)

r_R <- r_all_sliding_70_120 %>%
  mutate(measure = "r",
         value = r,
         reference = 0) %>%
  bind_rows(R)

fig2_3_70_120 <- r_R %>%
  ggplot(aes(x = w_max, y = value)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(aes(yintercept = reference), linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0,0, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside"
  ) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "", y = "") +
  scale_colour_manual(values = pal) +
  facet_grid(rows = vars(measure),
             scales = "free_y",
             switch = "y",
             labeller = as_labeller(c(r = "Daily growth rate (r)",
                                      R = "Effective reproduction\nnumber (Re)")))  

```

We combine the estimated growth rates and effective reproduction numbers into
a single figure.

```{r fig_2_3_all_ages, fig.height = 12}

ggpubr::ggarrange(fig2_3_0_18,
                  fig2_3_19_69,
                  fig2_3_70_120,
                  nrow = 3,
                  labels = "AUTO",
                  common.legend = TRUE,
                  legend = "bottom",
                  align = "hv") 


```



<!-- =======================================================  -->
<!-- =======================================================  -->
<!-- ======================================================= -->


# Comparison with deaths time series {.tabset .tabset-fade .tabset-pills}

## Outline

We want to explore the correlation between NHS Pathways reports and deaths, and
assess the potential for reports to be used as an early warning system for
disease resurgence.

Death data are publically available. We truncate the time series to
avoid bias from reporting delay - we assume a conservative delay of three weeks.


## Lagged correlation

We calculate Pearson's correlation coefficient between deaths and NHS Pathways
notifications using different lags. Confidence intervals are obtained using
bootstrap. Note that results were also confirmed using Spearman's rank
correlation.

First we join the NHS Pathways and death data, and aggregate over all England:

```{r data_agg}
## truncate death data for reporting delay
delay_max <- 21
trunc_date <- max(dth$date_report) - delay_max

dth_trunc <- dth %>%
  rename(date = date_report) %>%
  filter(date <= trunc_date) 

## join with notification data
all_data <- x %>% 
  filter(!is.na(nhs_region)) %>%
  group_by(date, nhs_region) %>%
  summarise(count = sum(count, na.rm = T)) %>%
  ungroup %>%
  inner_join(dth_trunc,
             by = c("date","nhs_region"))

all_tot <- all_data %>%
  group_by(date) %>%
  summarise(count = sum(count, na.rm = TRUE),
            deaths = sum(deaths, na.rm = TRUE)) 

```


We calculate correlation with lagged NHS Pathways reports from 0 to 30 days behind deaths:

```{r lag_cor}
## Function to calculate Pearson's correlation between deaths and lagged
## reports. Note that `pearson` can be replaced with `spearman` for rank
## correlation.

getcor <- function(x, ndx) {
  return(cor(x$deaths[ndx],
             x$note_lag[ndx],
             use = "complete.obs",
             method = "pearson"))
}

## Catch if sample size throws an error
getcor2 <- possibly(getcor, otherwise = NA)

getboot <- function(x) {
  result <- boot::boot.ci(boot::boot(x, getcor2, R = 1000), 
                          type = "bca")
  return(c(result$t0,result$bca[4:5]))
}



## Calculate all correlations + bootstrap CIs
lag_cor <- data.frame()
for (i in 0:30) {
  
  boot <- NULL
  boot_ci <- NULL
  
  ## lag reports
  temp <- all_tot %>% 
    mutate(note_lag = lag(count, i)) 
  
  ## calculate bootstrap CI for rank correlation
  boot <- getboot(temp)
  
  temp %>%
    summarise(r = boot[1],
              r_low = boot[2],
              r_hi = boot[3],
              lag = i,
              n = sum(!is.na(note_lag) & !is.na(deaths))) %>%
    select(lag, r,r_low, r_hi, n) -> summary
  
  lag_cor <- bind_rows(lag_cor, summary)
}

cor_vs_lag <- ggplot(lag_cor, aes(lag, r)) +
  theme_bw() +
  geom_ribbon(aes(ymin = r_low, ymax = r_hi), alpha = 0.2) +
  geom_hline(yintercept = 0, lty = "longdash") +
  geom_point() +
  geom_line() +
  labs(x = "Lag between NHS pathways and death data (days)",
       y = "Pearson's correlation") +
  large_txt
cor_vs_lag

l_opt <- which.max(lag_cor$r)

dth_vs_note <- all_tot %>% 
  mutate(note_lag = lag(count, lag_cor$lag[l_opt])) %>%
  filter(!is.na(note_lag)) %>%
  ggplot(aes(y = deaths, x = note_lag)) +
  geom_point() + 
  geom_smooth(color = "black",
              method = "glm", formula = "y ~ x") +
  theme_bw() +
  labs(y = "Daily number of \ndeaths reported",
       x = "Daily number of NHS Pathways reports") +
  large_txt
dth_vs_note

```



# Supplementary figures {.tabset .tabset-fade .tabset-pills}

## Serial interval distribution

This is a comparison of gamma versus lognormal distribution for the serial
interval used to convert r to R in our analysis. Both distributions are
parameterised with mean 4.7 and standard deviation 2.9.

```{r fig_serialinterval, fig.width = 10}
SI_param <- epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale, w = 0.5)

SI_distribution_alt <- distcrete::distcrete("lnorm", interval = 1,
                                            meanlog = log(4.7),
                                            sdlog = log(2.9), w = 0.5)

SI_dist1 <- as.data.frame(table(SI_distribution$r(1e5))) %>%
  ggplot(.) + 
  geom_col(aes(factor(Var1), Freq)) + 
  labs(x = "Serial interval (days)", y = "Frequency") +
  scale_x_discrete(breaks = seq(0, 100, 5)) +
  theme_bw()
SI_dist2 <- as.data.frame(table(SI_distribution_alt$r(1e5))) %>%
  ggplot(.) + 
  geom_col(aes(factor(Var1), Freq)) + 
  labs(x = "Serial interval (days)", y = "Frequency") +
  scale_x_discrete(breaks = seq(0, 500, 20)) +
  theme_bw()

ggpubr::ggarrange(SI_dist1,
                  SI_dist2,
                  nrow = 1,
                  labels = "AUTO") 


```


## Sensitivity analysis - 7 or 21 days moving window

We reproduce the window analysis with either a 7 or 21 days window for sensitivity
purposes.

First with the 7 days window:

```{r sliding_model_7}
## set moving time window (1/2/3 weeks)
w <- 7

# create empty df
r_all_sliding <- NULL

## make data for model
x_model_all_moving <- x %>%
  filter(!is.na(nhs_region)) %>% 
  group_by(date, nhs_region) %>%
  summarise(n = sum(count)) 

unique_dates <- unique(x_model_all_moving$date)

for (i in 1:(length(unique_dates) - w)) {
  
  date_i <- unique_dates[i]
  
  date_i_max <- date_i + w
  
  model_data <- x_model_all_moving %>%
    filter(date >= date_i & date < date_i_max) %>%
    mutate(day = as.integer(date - date_i)) %>% 
    day_of_week()
  
  
  mod <- glm(n ~ day * nhs_region + day_of_week,
             data = model_data,
             family = 'quasipoisson')
  
  # get growth rate
  r <- get_r(mod)
  r$w_min <- date_i
  r$w_max <- date_i_max
  
  # combine all estimates
  r_all_sliding <- bind_rows(r_all_sliding, r)
  
}

#serial interval distribution
SI_param = epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale,
                                        w = 0.5)

#convert growth rates r to R0
r_all_sliding <- r_all_sliding %>%
  mutate(R = epitrix::r2R0(r, SI_distribution),
         R_lower_95 = epitrix::r2R0(lower_95, SI_distribution),
         R_upper_95 = epitrix::r2R0(upper_95, SI_distribution))

```

```{r all_sliding_growth_7}
# plot
plot_growth <-
  r_all_sliding %>%
  ggplot(aes(x = w_max, y = r)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(colour = guide_legend(title = "",override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated daily growth rate (r)") +
  scale_colour_manual(values = pal)

```

```{r all_sliding_R_7}
plot_R <- r_all_sliding %>%
  ggplot(aes(x = w_max, y = R)) +
  geom_ribbon(aes(ymin = R_lower_95, ymax = R_upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated effective reproduction\nnumber (Re)") +
  scale_colour_manual(values = pal)

```

```{r fig2_3_7}

R <- r_all_sliding %>%
  mutate(lower_95 = R_lower_95, 
         upper_95 = R_upper_95,
         value = R,
         measure = "R",
         reference = 1)

r_R <- r_all_sliding %>%
  mutate(measure = "r",
         value = r,
         reference = 0) %>%
  bind_rows(R)

r_R_7 <- r_R %>%
  ggplot(aes(x = w_max, y = value)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(aes(yintercept = reference), linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0,0, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside"
  ) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "", y = "") +
  scale_colour_manual(values = pal) +
  facet_grid(rows = vars(measure),
             scales = "free_y",
             switch = "y",
             labeller = as_labeller(c(r = "Daily growth rate (r)",
                                      R = "Effective reproduction\nnumber (Re)")))


```

Then with the 21 days window:

```{r sliding_model_21}
## set moving time window (1/2/3 weeks)
w <- 21

# create empty df
r_all_sliding <- NULL

## make data for model
x_model_all_moving <- x %>%
  filter(!is.na(nhs_region)) %>% 
  group_by(date, nhs_region) %>%
  summarise(n = sum(count)) 

unique_dates <- unique(x_model_all_moving$date)

for (i in 1:(length(unique_dates) - w)) {
  
  date_i <- unique_dates[i]
  
  date_i_max <- date_i + w
  
  model_data <- x_model_all_moving %>%
    filter(date >= date_i & date < date_i_max) %>%
    mutate(day = as.integer(date - date_i)) %>% 
    day_of_week()
  
  
  mod <- glm(n ~ day * nhs_region + day_of_week,
             data = model_data,
             family = 'quasipoisson')
  
  # get growth rate
  r <- get_r(mod)
  r$w_min <- date_i
  r$w_max <- date_i_max
  
  # combine all estimates
  r_all_sliding <- bind_rows(r_all_sliding, r)
  
}

#serial interval distribution
SI_param = epitrix::gamma_mucv2shapescale(4.7, 2.9/4.7)
SI_distribution <- distcrete::distcrete("gamma", interval = 1,
                                        shape = SI_param$shape,
                                        scale = SI_param$scale,
                                        w = 0.5)

#convert growth rates r to R0
r_all_sliding <- r_all_sliding %>%
  mutate(R = epitrix::r2R0(r, SI_distribution),
         R_lower_95 = epitrix::r2R0(lower_95, SI_distribution),
         R_upper_95 = epitrix::r2R0(upper_95, SI_distribution))

```

```{r all_sliding_growth_21}
# plot
plot_growth <-
  r_all_sliding %>%
  ggplot(aes(x = w_max, y = r)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(colour = guide_legend(title = "",override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated daily growth rate (r)") +
  scale_colour_manual(values = pal)

```

```{r all_sliding_R_21}
# plot
plot_R <-
  r_all_sliding %>%
  ggplot(aes(x = w_max, y = R)) +
  geom_ribbon(aes(ymin = R_lower_95, ymax = R_upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0.5,0.5, "cm")) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "",
       y = "Estimated effective reproduction\nnumber (Re)") +
  scale_colour_manual(values = pal)

```

```{r fig2_3_21}

R <- r_all_sliding %>%
  mutate(lower_95 = R_lower_95, 
         upper_95 = R_upper_95,
         value = R,
         measure = "R",
         reference = 1)

r_R <- r_all_sliding %>%
  mutate(measure = "r",
         value = r,
         reference = 0) %>%
  bind_rows(R)

r_R_21 <- r_R %>%
  ggplot(aes(x = w_max, y = value)) +
  geom_ribbon(aes(ymin = lower_95, ymax = upper_95, fill = nhs_region), alpha = 0.1) +
  geom_line(aes(colour = nhs_region)) +
  geom_point(aes(colour = nhs_region)) +
  geom_hline(aes(yintercept = reference), linetype = "dashed") +
  theme_bw() +
  scale_weeks +
  theme(legend.position = "bottom",
        plot.margin = margin(0.5,1,0,0, "cm"),
        strip.background = element_blank(),
        strip.placement = "outside"
  ) +
  guides(color = guide_legend(title = "", override.aes = list(fill = NA)), fill = FALSE) +
  labs(x = "", y = "") +
  scale_colour_manual(values = pal) +
  facet_grid(rows = vars(measure),
             scales = "free_y",
             switch = "y",
             labeller = as_labeller(c(r = "Daily growth rate (r)",
                                      R = "Effective reproduction\nnumber (Re)")))

```

And we combine both outputs into a single plot:

```{r fig_sensitivity_window, fig.height = 8} 

ggpubr::ggarrange(r_R_7,
                  r_R_21,
                  nrow = 2,
                  labels = "AUTO",
                  common.legend = TRUE,
                  legend = "bottom") 

```



<!-- ======================================================= -->
<!-- ======================================================= -->
<!-- ======================================================= -->


# Export data {.tabset .tabset-fade .tabset-pills}

We save the tables created during our analysis:

```{r exports}

if (!dir.exists("excel_tables")) {
  dir.create("excel_tables")
}


## list all tables, and loop over export
tables_to_export <- c("r_all_sliding")

for (e in tables_to_export) {
  rio::export(get(e),
              file.path("excel_tables",
                        paste0(e, ".xlsx")))
}

```








<!-- =======================================================  -->
<!-- =======================================================  -->
<!-- ======================================================= -->

# System information {.tabset .tabset-fade .tabset-pills}

## Outline

The following information documents the system on which the document was
compiled.


## System 

This provides information on the operating system.

```{r system_info}
Sys.info()
```

## R environment

This provides information on the version of R used:

```{r R_session}
R.version
```


## R packages

This provides information on the packages used:

```{r R_pkg}
sessionInfo()
```
